---
output: github_document
---

<!-- 
README.md is generated from README.Rmd. Please edit that file
#knitr::knit("README.Rmd") 
rmarkdown::render("README.Rmd") 
maybe clear cache before
-->


```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```


[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/REddyProc)](http://cran.r-project.org/package=REddyProc)
[![Travis-CI Build Status](https://travis-ci.org/bgctw/REddyProc.svg?branch=master)](https://travis-ci.org/bgctw/REddyProc)


## Overview

`REddyProc` package supports processing (half)hourly data from Eddy-Covariance sensors.

There is an online-formular to use the functionality of the package including
description at
<https://www.bgc-jena.mpg.de/bgi/index.php/Services/REddyProcWeb>.

## Installation
```{r, eval = FALSE}
# Release stable version from CRAN
install.packages("REddyProc")

# The development version from GitHub using devtools:
# install.packages("devtools")
devtools::install_github("bgctw/REddyProc")
```

## Usage

A simple example performs Lookuptable-based gapFilling of
Net-Ecosystem-Exchange (NEE) and plotting a fingerprint plot of the filled
values.

```{r example, results='hide', message=FALSE, cache=TRUE}
library(REddyProc)
#+++ Input data from csv (example needs to be downloaded)
examplePath <- getExamplePath('Example_DETha98.txt', isTryDownload = TRUE)
if (length(examplePath)) {
  EddyData.F <- fLoadTXTIntoDataframe(examplePath)
} else {
  warning(
      "Could not find example text data file."
      ," In order to execute this example code,"
      ," please, allow downloading it from github. " 
      ," Type '?getExamplePath' for more information.")
  # using RData version distributed with the package instead
  EddyData.F <- Example_DETha98
}
#+++ If not provided, calculate VPD from Tair and rH
EddyData.F$VPD <- fCalcVPDfromRHandTair(EddyData.F$rH, EddyData.F$Tair)
#+++ Add time stamp in POSIX time format
EddyDataWithPosix.F <- fConvertTimeToPosix(
  EddyData.F, 'YDH', Year.s = 'Year', Day.s = 'DoY', Hour.s = 'Hour')
#+++ Initalize R5 reference class sEddyProc for processing of eddy data
#+++ with all variables needed for processing later
EddyProc.C <- sEddyProc$new(
  'DE-Tha', EddyDataWithPosix.F, c('NEE','Rg','Tair','VPD', 'Ustar'))
#Location of DE-Tharandt
EddyProc.C$sSetLocationInfo(
  Lat_deg.n = 51.0, Long_deg.n = 13.6, TimeZone_h.n = 1)  
#
#++ Fill NEE gaps with MDS gap filling algorithm (without prior ustar filtering)
EddyProc.C$sMDSGapFill('NEE', FillAll.b = FALSE)
#
#++ Export gap filled and partitioned data to standard data frame
FilledEddyData.F <- EddyProc.C$sExportResults()
#
#++ Example plots of filled data to screen or to directory \plots
EddyProc.C$sPlotFingerprintY('NEE_f', Year.i = 1998)
```



Further examples are in
[vignette(useCase)](https://github.com/bgctw/REddyProc/blob/master/vignettes/useCase.md)
and
[vignette(DEGEbExample)](https://github.com/bgctw/REddyProc/blob/master/vignettes/DEGebExample.md)


## Docker images
Docker images are provided that comprise rstudio, rocker/tidyverse, and REddyProc.
There are different version for the latest push to github, for the version on CRAN and for specific tags starting 
from 1.1.4.

- bgctw/reddyproc:latest   
- bgctw/reddyproc_cran
- bgctw/reddyproc:`tag`

They are usually run with installed docker by saying at a shell:
```
docker run --rm -p 8787:8787 <imagename>
```
Then the url `localhost:8787` should bring up RStudio in the browser window, where
you can type the above usage example.




