---
output: 
  rmarkdown::html_vignette:
    keep_md: true
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{different treatments of uStar threshold}
  %\usepackage[UTF-8]{inputenc}

---

```{r, include = FALSE}
# do not execute on CRAN: 
# https://stackoverflow.com/questions/28961431/computationally-heavy-r-vignettes
is_check <- ("CheckExEnv" %in% search()) || any(c("_R_CHECK_TIMINGS_",
             "_R_CHECK_LICENSE_") %in% names(Sys.getenv()))
knitr::opts_chunk$set(eval = !is_check)
```

```{r setup, include = FALSE}
library(knitr)
#rmarkdown::render("vignettes/uStarCases.Rmd","md_document")
opts_knit$set(root.dir = '..')
opts_chunk$set(
    #, fig.align = "center"
    #, fig.width = 3.27, fig.height = 2.5, dev.args = list(pointsize = 10)
    #,cache = TRUE
    #, fig.width = 4.3, fig.height = 3.2, dev.args = list(pointsize = 10)
    #, fig.width = 6.3, fig.height = 6.2, dev.args = list(pointsize = 10)
    # works with html but causes problems with latex
    #,out.extra = 'style = "display:block; margin: auto"' 
    )
knit_hooks$set(spar = function(before, options, envir) {
    if (before) {
        par(las = 1 )                   #also y axis labels horizontal
        par(mar = c(2.0,3.3,0,0) + 0.3 )  #margins
        par(tck = 0.02 )                          #axe-tick length inside plots             
        par(mgp = c(1.1,0.2,0) )  #positioning of axis title, axis labels, axis
     }
})
```


```{r, include = FALSE}
#themeTw <- theme_bw(base_size = 10) + theme(axis.title = element_text(size = 9))
```

# Different treatments of uStar threshold

The recommended way of dealing with the uncertain uStar threshold for filtering
the half-hourly data, is to repeat all the processing steps with several 
bootstrapped estimates of the threshold as in `vignette('useCase')`.

First, some setup.

```{r init, message = FALSE, output = 'hide'}
#+++ load libraries used in this vignette
library(REddyProc)
library(dplyr)
#+++ define directory for outputs
outDir <- tempdir()  # CRAN policy dictates to write only to this dir in examples
#outDir <- "out"     # to write to subdirectory of current users dir
#+++ Add time stamp in POSIX time format to example data 
# and filter long runs of equal NEE values
EddyDataWithPosix <- fConvertTimeToPosix(
  filterLongRuns(Example_DETha98, "NEE")
  , 'YDH', Year = 'Year', Day = 'DoY', Hour = 'Hour')
```

## Not applying uStar filtering

Subsequent processing steps can be performed without further uStar filtering
using `sEddyProc_sMDSGapFill`. Corresponding result columns then have
no uStar specific suffix. 

```{r noUStar, message = FALSE}
EProc <- sEddyProc$new(
  'DE-Tha', EddyDataWithPosix, c('NEE','Rg','Tair','VPD', 'Ustar'))
EProc$sMDSGapFill('NEE')
grep("NEE.*_f$",names(EProc$sExportResults()), value = TRUE)
```

## User-specified uStar threshold

The user can provide value for uStar-filtering before gapfilling, using
`sEddyProc_sMDSGapFillAfterUstar`. Output columns for this uStar scenario use
the suffix as specified by argument `uStarSuffix` which defaults to "uStar". 

The friction velocity, uStar, needs to be in column named "Ustar" of the input 
dataset.

```{r userUStar, message = FALSE}
EProc <- sEddyProc$new(
  'DE-Tha', EddyDataWithPosix, c('NEE','Rg','Tair','VPD', 'Ustar'))
uStar <- 0.46
EProc$sMDSGapFillAfterUstar('NEE', uStarTh = uStar)
grep("NEE.*_f$",names(EProc$sExportResults()), value = TRUE)
```

## Single uStar threshold estimate

The uStar threshold can be estimated from the uStar-NEE relationship 
from the data without estimating its uncertainty by a bootstrap.

```{r singleUStar, message = FALSE}
EProc <- sEddyProc$new(
  'DE-Tha', EddyDataWithPosix, c('NEE','Rg','Tair','VPD', 'Ustar'))
# estimating the thresholds based on the data (without bootstrap)
(uStarTh <- EProc$sEstUstarThold())
# may plot saturation of NEE with UStar for a specified season to pdf
EProc$sPlotNEEVersusUStarForSeason(levels(uStarTh$season)[3], dir = outDir )
```

Next, the annual estimate is used as the default in gap-filling.
Output columns use the suffix as specified by argument `uSstarSuffix` 
which defaults to "uStar".

```{r singleUStarGapfill, message = FALSE}
#EProc$useAnnualUStarThresholds()
EProc$sMDSGapFillAfterUstar('NEE')
grep("NEE.*_f$",names(EProc$sExportResults()), value = TRUE)
```

## Scenarios across distribution of u* threshold estimate

Choosing a different u* threshold effects filtering and the subsequent processing
steps of gap-filling, and flux-partitioning. In order to quantify the uncertainty
due to not exactly knowing the u* threshold, these processing steps should be
repeated for different threshold scenarios, and the spread across the results should
be investigated.

First, the quantiles of the threshold distribution are estimated by bootstrap.

```{r uStarScen, results='hold'}
EProc <- sEddyProc$new(
  'DE-Tha', EddyDataWithPosix, c('NEE','Rg','Tair','VPD', 'Ustar'))
EProc$sEstimateUstarScenarios(
    nSample = 100L, probs = c(0.05, 0.5, 0.95))
# inspect the thresholds to be used by default
EProc$sGetUstarScenarios()
```

By default the annually aggregated threshold estimates are used for each season
within one year as in the original method publication.
To see the estimates for different aggregation levels,
use method `sEddyProc_sGetEstimatedUstarThresholdDistribution`:
```{r}
(uStarThAgg <- EProc$sGetEstimatedUstarThresholdDistribution())
```

In conjunction with method `usGetSeasonalSeasonUStarMap` and 
`sEddyProc_sSetUstarScenarios` this can be used
to set seasonally different u* threshold. 
However, this common case supported by method 
`sEddyProc_useSeaonsalUStarThresholds`. 

```{r uStarScenSetSeasonal}
#EProc$sSetUstarScenarios(
#  usGetSeasonalSeasonUStarMap(uStarThAgg)[,-2])
EProc$useSeaonsalUStarThresholds()
# inspect the changed thresholds to be used
EProc$sGetUstarScenarios()
```

Several function whose name ends with 'UstarScens'
perform the subsequent processing steps for all uStar scenarios.
They operate and create columns that differ between threshold scenarios by
a suffix.

```{r uStarScenGapfill, message=FALSE}
EProc$sMDSGapFillUStarScens("NEE")
grep("NEE_.*_f$",names(EProc$sExportResults()), value = TRUE)
```

```{r uStarScenMRPart, message=FALSE}
EProc$sSetLocationInfo(LatDeg = 51.0, LongDeg = 13.6, TimeZoneHour = 1)
EProc$sMDSGapFill('Tair', FillAll = FALSE, minNWarnRunLength = NA)
EProc$sMDSGapFill('Rg', FillAll = FALSE, minNWarnRunLength = NA)
EProc$sMDSGapFill('VPD', FillAll = FALSE, minNWarnRunLength = NA)
EProc$sMRFluxPartitionUStarScens()
grep("GPP_.*_f$",names(EProc$sExportResults()), value = TRUE)
if (FALSE) {
  # run only interactively, because it takes long
  EProc$sGLFluxPartitionUStarScens(uStarScenKeep = "U50")
  grep("GPP_DT_.*_f$",names(EProc$sExportResults()), value = TRUE)
}
```

The argument `uStarScenKeep = "U50"` specifies that the outputs that
are not distinguished by the suffix, e.g. `FP_GPP2000`, should be reported for the
median u* threshold scenario with suffix `U50`, instead of the default first scenario.

## See also
A more advanced case of user-specified seasons for
uStar threshold estimate is given in [`vignette('DEGebExample')`](DEGebExample.html).
